### 2021-05-27 scripts/ZhangetalReanalysis.Rmd
updating paper

---
title: "Forest cover runoff relationship"
author: "Willem Vervoort, Eliana Nervi, Jimena Alonso"
date: "01/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(lubridate)
require(mgcv) # to do the generalised modelling
```

## Introduction

There has been an long and on-going discussion in the hydrologcal literature around the impact of forests on streamflow [@andreassian2004; @jackson2005; @zhang2017; @brown2005; @brown2013]. The historic work highlights provides a general consensus that if forest areas increase, streamflow decreases and vice-versa. The most dramatic result in relation to this, is Figure 5 in @zhang2011 indicating (or Australian watersheds) a 100% decrease in stream flow for watersheds with 100% forest cover. However, on the other end of the spectrum, in a series of French watersheds [@cosandey2005], there was no change in streamflow characteristics in 2 of the three watersheds studied in relation to deforestation.

There have been several review papers aiming to summarize different studies across the globe, in relation to paired watershed studies [@brown2005; @hewlett1982] and more generally [@jackson2005; @zhang2017]. These studies are aiming to generalize the individual findings and to identify if there are global trends or relationships that can be developed. The most recent review [@zhang2017] developed an impressive database of watershed studies in relation to changes in streamflow due to changes in forest cover based on a global data set. This dataset, that covers over 250 studies are described in terms of the change in streamflow as a result of the change in forest cover, where studies related to both forestation (increase in forest cover) and deforestation (decrease in forest cover) were included.

The conclusions of the paper [@zhang2017] suggest that there is a distinct difference in the change in flow as a result of forestation or deforestation between small watersheds, defined as < 1000 km^2^ and large watersheds > 1000 km^2^. While for small watersheds there was no real change in runoff with changes in cover, for large watersheds there was a clear trend showing a decrease in runoff with and increase in forest cover. Their main conclusion was that the response in annual runoff to forest cover was scale dependent and appeared to be more sensitive to forest cover change in water limited watersheds relative to energy limited watershed [@zhang2017].

Encouraged by the work presented by @zhang2017 and the fantastic database of studies presented by these authors, we believe we can add to the discussion by presenting further analysis of the data and by adding further watersheds and enhancements to the data base.

In particular, the main method in the work by @zhang2017 is using simple linear regression. And the main assumption is that the threshold at 1000 km^2^ is a distinct separation between "small" and "large" watersheds. Given the fantastic data set collected, the analysis can be easily expanded to look at interactions between the terms and to test the assumption of a distinct threshold at 1000 km^2^.

In particular, the objective of this paper is to 1) enhance the data set from @Zhang2017 with further watersheds and spatial coordinates and 2) to analyse the possibility of non-linear and partial effects of the different factors and variables in the data base using generalised linear (GLM) and generalised additive models (GAM @wood2006). Finally we hope to point to further research that can expand our work and that outlined @zhang2017 to better understand the impact of forest cover change on streamflow.

## The data

The starting point consists of the two tables in the supplementary data from Zhang et al. (2017), which means there is one table for small watersheds (< 1000 km2) and one table for large watersheds (> 1000 km2). The columns in the data represent We might combine these tables as we can easily add a variable indicating whether the watershed originates from the small or large table.

```{r}
Zhang_small <- read_csv("TablesZhangetal2017_small.csv")
Zhang_large <- read_csv("TablesZhangetal2017_large.csv")
```

### change names of columns

the package`mgcv` does not seem to be happy with complex names, so probably better to make all names simple

```{r}
names(Zhang_small)[3:4] <- c("Area_km2", "Pa_mm")
names(Zhang_large)[3:4] <- c("Area_km2", "Pa_mm")
names(Zhang_small)[5:6] <- c("Forest_type", "Hydrological_regime")
names(Zhang_large)[5:6] <- c("Forest_type", "Hydrological_regime")
names(Zhang_small)[9:10] <- c("Precip_data_type", "Assessment_technique")
names(Zhang_large)[9:10] <- c("Precip_data_type", "Assessment_technique")
```

## Simple generalised linear model

This will use the pacakge `mgcv`, which can do both generalised linear modelling and generalised additive modelling

We will start with a simple linear model which essentially predicts the change in streamflow $\Delta Qf\%$ as a function of the other variables:
$\Delta Qf\% ~ Area + Pa + Forest type + Hydrological regime + \Delta F\% + Precipitation data type + Assessment technique$

### Small watersheds

```{r}
linear_model_small <- gam(DeltaQf_perc ~ DeltaF_perc + Area_km2 + Pa_mm + Forest_type + 
                            Hydrological_regime + Precip_data_type +  Assessment_technique, data = Zhang_small)
summary(linear_model_small)
gam.check(linear_model_small)

```
This shows that the change in forestry is a highly signficant predictor of the change in flow, and this indicates a decrease in the change in flow withe an increase in the change in forestry. In other words it shows the same relationship as figure 2 in Zhang et al.
However, it is dependent on rainfall (with higher rainfall showing steeper declines) and the type of forest (with Mixed forest types causing a greater decline in flow with an increase in forestry). None of the other variables are significant.

The residuals of the model are slightly problematic, they are skewed and show some increase in variance. This is probably due to the distribution of the y-variable (the change in flow).

Check distribution of the y-variable

```{r}
hist(Zhang_small$DeltaQf_perc)
min(Zhang_small$DeltaQf_perc)
```
Doing a log transformation on the y-variable is problematic as the change in flow ranges from -200% to +200%. However, it is possible when you divide the percentage by 100, and then use a log10(y + 1) transformation.

```{r}
linear_model_small_log <- gam(log10(DeltaQf_perc/100 + 1) ~ DeltaF_perc + Area_km2 + Pa_mm + Forest_type + 
                            Hydrological_regime + Precip_data_type +  Assessment_technique, data = Zhang_small)
summary(linear_model_small_log)
gam.check(linear_model_small_log)
```
This improves the residuals, but the results of the analysis is the same.

### Large watersheds

```{r}
linear_model_large <- gam(DeltaQf_perc ~ DeltaF_perc + Area_km2 + Pa_mm + Forest_type + Hydrological_regime + Precip_data_type +
                    Assessment_technique, data = Zhang_large)
summary(linear_model_large)
gam.check(linear_model_large)
```
The results for the last watersheds shows that in this case the change in forestry is not significant at p < 0.05 as the p-value is 0.11. Increasing Rainfall is associated with a lower change in runoff, while having a snow dominated hydrological regime is associated with a higher change in runoff (at p <0.05). The "SH" assessment technique (combined use of hydrographs and statistical techniques) is also associated with a lower change in runoff.

The residuals are again a little skewed, again because the change in runoff percentage data is skewed.

```{r}
hist(Zhang_large$DeltaQf_perc)
```

Using a similar transformation as for the small watersheds results on the change in streamflow, and a log transformation of the rainfall however results in quite a different model outcome

```{r}
linear_model_large_log <- gam(log10(DeltaQf_perc/100+1) ~ DeltaF_perc + Area_km2 + log10(Pa_mm) + Forest_type + Hydrological_regime + Precip_data_type +
                    Assessment_technique, data = Zhang_large)
summary(linear_model_large_log)
gam.check(linear_model_large_log)

```
The residuals are much better distributed. And the results now show that the change in forestry is a significant predictor, resulting in a decrease of the change in stream flow with an increase in the change in forestry.. In addition, increasing rainfall also decreases the change in streamflow. while snow dominated  regimes, hydrological modelling techniques and statistical and hydrograph approaches predict higher changes in streamflow.

## is the separation in large and small arbitrary

Stack the two data frames and run the GAM and check if the size of the watershed is significant.

```{r}
Zhang_all <- bind_rows(Zhang_small, Zhang_large)

linear_model_all_log <- gam(log10(DeltaQf_perc/100+1) ~ DeltaF_perc + Area_km2 + log10(Pa_mm) + Forest_type + Hydrological_regime + Precip_data_type +
                    Assessment_technique, data = Zhang_all)
summary(linear_model_all_log)
gam.check(linear_model_all_log)

```

## Work to be done

1. Extend the analysis to include a smooth non-linear rainfall response?  
2. In the paper, they also use PET which is not in the data, se we need to find some value of PET for each watershed. I wonder how you define PET for the whole of the Amazon?
3. We also probably want to extract the "continent" and I think also Latitude  
4. We also need to check if we know of more published data that is not in the current dataset

Key arguments we are making are:  
1. regression approach is too simple, ignores possible interactions and possible non-linear behaviour (for example rainfall).  
2. Is the 1000 km2 is an arbitrary cut-off?  Or is there an interaction with the assessment technique?  
3. Can you actually define "annual rainfall" or "annual PET" for a large watershed? Would there not be a major effect of spatial variability?
4. time period after which measurement of decrease in flow is reached: harvesting, growth cycles etc. There is a need for standardisation or incorporation of this variability.