### 2022-08-09 paper/Forest_and_Water/SupplementaryMaterialPart2.Rmd

---
title: 'Supplementary Information part 2: Testing the improved data sets'
author: "Willem Vervoort/Eliana Nervi/Jimena Alonso"
date: "`r Sys.Date()`"
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output:   
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: forestandwater.bib
editor_options: 
  chunk_output_type: inline
header-includes:
  - \usepackage{setspace}
  - \usepackage{color}
abstract: |
  This supplementary material file compares whether the inclusion of additional catchments generates fundamentally different results as the original (but improved) data. Single variable regressions on the smaller (original) dataset are compared with the extended data set.
---

# Introduction

In this document we tested whether the fundamental conclusions in the single variable regressions with the improved data base differed from the original conclusions in @zhang2017. This is to check how much influence the changes to the data set and the additional data might have changed the original conclusions.

First we will read in the data

```{r setup, include=FALSE, echo =T}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lubridate)
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))

Zhang_small <- read_csv("../../data/SmallCatchments_T2.csv") %>%
  mutate(table = "Small catchments")
Zhang_large <- read_csv("../../data/LargeCatchments_T1.csv") %>%
  mutate(table = "Large catchments")
new_data <- read_csv("../../data/NewCatchments_T3.csv") %>%
  mutate(table = "new catchments")

# change some column names
names(Zhang_small)[3:4] <- c("Area_km2", "Pa_mm")
names(Zhang_large)[3:4] <- c("Area_km2", "Pa_mm")
names(new_data)[3:4] <- c("Area_km2", "Pa_mm")
names(Zhang_small)[5:6] <- c("Forest_type", "Hydrological_regime")
names(Zhang_large)[5:6] <- c("Forest_type", "Hydrological_regime")
names(new_data)[5:6] <- c("Forest_type", "Hydrological_regime")
names(Zhang_small)[9:10] <- c("Precip_data_type", "Assessment_technique")
names(Zhang_large)[9:10] <- c("Precip_data_type", "Assessment_technique")
names(new_data)[9:10] <- c("Precip_data_type", "Assessment_technique")

# combine the data into one dataset

```

We will combine the different tables, but will keep an indicator to see where the data are from.

```{r, echo = T, warning = F, message = F}
Zhang_small$From <- as.numeric(Zhang_small$From)
Zhang_small$To <- as.numeric(Zhang_small$To)
Zhang_all <- bind_rows(Zhang_large,Zhang_small) %>%
  mutate(dataset = "original Zhang et al data")
new_data <- new_data %>%
  mutate(dataset = "new data")
All_data <- bind_rows(Zhang_all, new_data)
```

### Implementing the changes to the overall data

The following code implements the changes described in the Supplementary data part 1. However, many of the changes were implemented manually into the data set. These are simply the remaining changes not implemented manually.

1. removing the duplicates.  
```{r, echo = T}
All_data <- All_data %>%
  mutate(`Possible duplicate` = 
           ifelse(is.na(`Possible duplicate`)==T,0,`Possible duplicate`),
         `Possible duplicate` = as.numeric(`Possible duplicate`)) %>%
  filter(`Possible duplicate` != 1)
```

2. calculating the dryness

```{r, echo = T}
# calculate dryness index
All_data <- All_data %>%
  mutate(Dryness = E0/Pa_mm)
```

3. remove watershed 1 (the Amazon) from the analysis  
```{r, echo = T}
All_data <- All_data %>%
  filter(`Watershed #` != 1)
```

4. remove data set 188 and 254 Kamakia and Sambret
```{r,echo = T}
All_data <- All_data %>%
  filter(`Watershed #` != 188) %>%
  filter(`Watershed #` != 254)
```

5. add a column that indicates forst loss of forest gain

```{r, echo = T}
All_data <- All_data %>%
  mutate(forest_sign = ifelse(DeltaF_perc < 0, "Forest Cover Loss", "Forest Cover Gain"))
```

# Methods

The approach is similar to @zhang2017. We run single variable regressions separating large (> 1000 km^2^) and small catchments (<= 1000 km^2^).

The paper by @zhang2017 calculates the sensitivity of runoff as a function of runoff as:

$\Delta Q_f = 100 \times \frac{\Delta Q_{f,mm}}{\bar{Q}}$

This first equation is superfluous in this case as the data (as extracted from @zhang2017) is already defined in terms of $\Delta Q_f$.

$S_f = \frac{\Delta Q_f}{\Delta F}$

```{r, echo = T}
All_data <- All_data %>%
  filter(is.na(DeltaF_perc) == F) %>%
  mutate(S_f = DeltaQf_perc/DeltaF_perc)
```

In sequence we analyse:

- the relationship between forest cover change and streamflow change for small and large catchments (i.e. Figure 2 in @zhang2017);  
- the relationship between catchment size and the sensitivity to runoff change (i.e. Figure 3 in @zhang2017); and  
- the sensitivity to forest loss as a function of dryness (i.e. Figure 4 in @zhang2017).

# Results

1. The change in flow as a function of change in forest cover

First the results for the original (improved) dataset

```{r, echo = F, message = F}
png("Fig2Zhang.png", width = 2500, height = 1800, res = 300)
All_data %>%
  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

# and for all data sets
png("Fig2Zhang_all.png", width = 2500, height = 1800, res = 300)
All_data %>%
#  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign, shape = table)) + 
  geom_point() +  geom_smooth(method = "lm") +
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

```

```{r Fig2Zhang, echo =F, fig.cap="Changes in flow based on the catchments from the original data set", out.width="90%"}

include_graphics("Fig2Zhang.png")
```
And then the results for all the catchments including the new "added" catchments.

```{r Fig2Zhangnew, echo =F, fig.cap="Changes in flow based on the catchments from the extended data set", out.width="90%"}

include_graphics("Fig2Zhang_all.png")
```