### 2022-10-03 paper/Forest_and_Water/Test_model_totalForestArea.Rmd

---
title: "Testing Forest Area as a random effect"
author:
  - name: R. Willem Vervoort
    email: willem.vervoort@sydney.edu.au
    footnote: 1
  - name: Eliana Nervi
    email: eliananervif@gmail.com
  - name: Jimena Alonso
    email: jalonso@fing.edu.uy
footnote:
  - code: 1
    text: "Corresponding Author"    
date: "`r Sys.Date()`"
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output:   
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: forestandwater.bib
#editor_options: 
#  chunk_output_type: inline
header-includes:
  - \usepackage{setspace}
  - \usepackage{color}
  - \newcommand{\beginsupplement}{  \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}\setcounter{equation}{0} \renewcommand{\theequation}{S\arabic{equation}}}
abstract: |
  This supplementary material file compares different linear and non-linear relationships for the impact of forest cover on the change in stream flow.
---
\beginsupplement

# Introduction

This supplementary material is related to 'Generalising the impact of forest cover on streamflow from experimental data: it is not that simple. Vervoort et al.'

This document outlines a test suggested in the first review of the manuscript. The AE of the Journal of Hydrology suggested that we test if including the total forest area as a variable would make sense. Not only the \% change in area is important, but what total area of forest was there in the first place.

This test focusses on the large catchment data (< 1000 km^2^) for two reasons:

1. This was a manageable subset and many of the papers are fairly accessible, in contrast to the small catchments.
2. Many of the small catchments have 100\% area covered in forest, so are not useful to identify if total area of forest has an impact on the change in flow.

## set up the packages

```{r setup, include=FALSE, echo =T}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lubridate)
require(mgcv)
require(visreg)
require(patchwork)
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))
```

## read in the data

Read in the data and massage column names to make columns useable in the model.

```{r data-w-forest}

Data_wF <- read_csv("../../data/LCdatawithFA.xlsx - LW_TotForest_Area.csv")

# change some column names
names(Data_wF)[3:4] <- c("Area_km2", "Pa_mm")
names(Data_wF)[5:6] <- c("Forest_type", "Hydrological_regime")
names(Data_wF)[13:14] <- c("Precip_data_type", "Assessment_technique")

```

## Some changes to the overall data

1. calculating the dryness

```{r, echo = T}
# calculate dryness index
Data_wF <- Data_wF %>%
  mutate(Dryness = E0/Pa_mm)
```

2. remove watershed 1 (the Amazon) from the analysis  
```{r, echo = T}
Data_wF <- Data_wF %>%
  filter(`Watershed #` != 1)
```

3. include length as a variable

```{r length, echo =T}
All_data <- Data_wF %>%
  mutate(length = To - From,
         mid_year = From + (To - From)/2)

```

# Results

Run the model simply as a GLM, transform Area using log10, as in the manuscript.

```{r model-percent}

Forest_model_all <- gam(DeltaQf_perc ~ DeltaF_perc + 
                    log10(Area_km2) + 
                    Dryness + 
                    Perc_Farea_pre +
                    Precip_data_type +  Assessment_technique +
                    Forest_type +
                    Hydrological_regime
                    , data = All_data)
summary(Forest_model_all)
```
This suggest that forest area is not at all significant in the model. In fact, the hydrological modelling technique is much more important. Overall the model explains roughly 82% of the variation.

```{r check-model-percent}
gam.check(Forest_model_all)
#plot(Forest_model_all)
```
The fit of the model is also well-behaved, as would be expected. The data are mostly from hydrological modelling, stabilising the response of the change in forest to change in streamflow.

## run the model on forest area rather than \%

Possibly, using the total \% forest is not taking into account the differences in total area of forest. So we can also test with total ha. Note that this can interact with total area: larger catchments will have larger total areas of forest.

We appear to have less data here as \% is reported more, but we can calculate total ha by multiplying \% by total area.

```{r model-area}
All_data <- Data_wF %>%
  mutate(Farea_km2 = ifelse(is.na(Farea_km2)==T,Area_km2*Perc_Farea_pre/100,Farea_km2))

#write_csv(All_data,"../../data/Data_TotForest_Area.csv")

Forest_model_area <- gam(DeltaQf_perc ~ DeltaF_perc + 
                    log10(Area_km2) + 
                    Dryness + 
                    Farea_km2 +
                    Precip_data_type +  Assessment_technique +
                    Forest_type +
                    Hydrological_regime
                    , data = All_data)
summary(Forest_model_area)
```
```{r check_resid_area}
gam.check(Forest_model_area)
#plot(Forest_model_all)
```

Again, not significant and the model is well-behaved. Note that the Hydrological modelling assessment technique (HM) is again highly significant.

# Conclusion

There is no effect of forest area in the relationship between change in flow  and percent change in forest cover.